### 2015-07-04 narrative/placement.tex
started on the data placement section

\subsection{Data Placement and Movement}

The application workflows targeted in this proposal generate very large
amounts of data, and this data must be processed and analyzed before
potential insights from the simulations can be realized. In-situ/in-transit
data processing using in-memory data-staging approaches have been
effectively used to address data-related challenges and enable simulations
workflows on current high end systems -- the key idea is to use compute and
storage resources on the high-end system to process the data close to where
it is generated. However, given current architectural trends, indicate that
emerging systems will have increasing numbers of cores per node and
correspondingly decreasing amounts of DDRAM memory per core as well as
decreasing memory bandwidth. Architectural trends also indicate deeper
memory hierarchies and multilevel storage structures. As a results, to
address the increases data volumes and data processing requirements requires
that these levels of storage be effectively utilized and exposes as a
uniform staging abstraction that can be used by the application workflows. A
key requirements is application-driven runtime mechanisms for dynamically
managing data placement across the layers of the distributed storage
hierarchy, coordinating data movement and data sharing between the
components of the application workflow so as to maximize its utility to the
application and reduce access costs.

@@

\paragraph{Proposed research approach:} In this project we will utilizes
deep memory hierarchies and multilevel storage structures to support
staging-based in-transit data management and processing for data-intensive
scientific workflows at extreme scale. Specifically, we will address the
coordination and data access/sharing requirements of application workflows,
as well as the online data management requirements such as for example,
in-transit online data analysis, check-pointing, data reduction. We will
also deal with the complexities associated with data placement of data and
movement in the multiple layers and components of the storage hierarchy. The
key components of our research include:

@@

\subsection{Migration}

@@

\subsection{Eviction Metrics}

>>>>

\subsection{Data Placement and Movement}
\newcommand{\Sir}{Sirocco}

The application workflows targeted in this proposal generate very large
amounts of data, which needs to be processed and analyzed before
potential insights from the simulations can be realized. 
In-situ and in-transit data staging approaches can alleviate some of the
burden on storage, leveraging available compute and memory resources on high
end systems to process the data close to where it is generated. 
However, given current architectural trends, emerging systems will have a
vastly great gulf between the computing capability (the capability to
generate data) and the storage capability (the capability to store the
generated data). Moreoever, the complexity and heterogeniety of the
architecture will introduce new tiers in the storage system. Instead of
being a flat, roughly homogenous space in which data can be stored, the new
storage hierarchy will be multi-tiered and greatly heterogeneous.

To take advantage of these architectural trends, we propose a triplet of
data placement strategies. First, we will utilize the faster layer of the
storage hierarhcies, such as DRAM, NVRAM and remote NVRAM, as the target for
the initial write from an application. To mitigate the space concerns, we
will use the techniques described in Section~\ref{sec:refactoring} to
identify and partition the segments of data with the highest utility for the
application and use case. The rest of the data will be stored on the
parallel file system, as is done today. Second, we will utilize the
migration of data supported by the existing implementation of \Sir to move the
data to an appropriate tier. We will explore how migration can be guided by
a combination of the utility function defined by the application and the
resources avialable to the storage system. Knowing the specific structural
and semantic information (including relationships between data segments)
about the data will enable migration to partition the data further to
optimize access to data for reader. 
Finally, we will explore policies that combine the knowledge of data, its
utility to users, the cost of maintaining data in long term storage and
other metrics to decide how to purge irrelevant data from the storage
system.

Thus we propose to advance the state of the art in three connected areas:
(1) Data placement,
(2) Migration,
and (3) Purging.

\subsubsection{Data Placement}
\label{sec:data-placement}

Our past work on data
staging\cite{tongipdps15,qiansc15}\cite{docan2012dataspaces}\cite{abbasi2010datastager}
in HPC systems, coupled with the experience in developing high performance
I/O transports for ADIOS \cite{liu2014hello,lofstead2008flexible} has
provided us with insights on how to manage the complex I/O requirements for
exascale applications, particularly in the face of the aforementioned
multi-tier storage hierarchy. A key requirements is application-driven
%
runtime mechanisms for dynamically managing data placement across the layers
of the distributed storage hierarchy,
%
coordinating data movement 
%
and data sharing between the components of the application workflow so as to
maximize its utility to the application and reduce access costs. Our
approach to data placement will leverage the increased set of knowledge
about the data available to the middleware and storage system, potentially
allowing for greater optimizations as compared to other storage solutions.

@@

\paragraph{Proposed research approach:} In this project we will utilize
deep memory hierarchies and multilevel storage structures to support
staging-based in-transit data management and processing for data-intensive
scientific workflows at extreme scale. Specifically, we will address the
coordination and data access/sharing requirements of application workflows,
as well as the online data management requirements such as for example,
in-transit online data analysis, check-pointing, data reduction. We will
also deal with the complexities associated with data placement of data and
movement in the multiple layers and components of the storage hierarchy. The
key components of our research include:

@@

\subsubsection{Migration}
\label{sec:migration}

@@

\subsubsection{Purging}
\label{sec:purging}
